---
title: ""
format: 
  docx:
    page-width: 10
    reference-doc: template.docx
editor: source
bibliography: resources/references.bib
citation-abbreviations: resources/medline-abbreviations.json
csl: resources/ama.csl
execute:
  echo: false
  warning: false
project: 
  execute-dir: project
crossref: 
  fig-title: Fig
filters:
  - resources/docx-landscape.lua
---

```{r render-for-colleagues, eval = FALSE, include = FALSE}
#' WHEN READY TO RENDER VERSION FOR SHARING WITH COLLEAGUES USE THIS CODE 
#' TO APPEND A DATE TO THE OUTPUT
save <- FALSE
if (save == TRUE) {
file_name <- paste0(Sys.Date(),"_draft_NPA-MEPS-comparison-manuscript.docx")

quarto::quarto_render('manuscript/QUARTO-Manuscript.qmd', output_file = file_name)

file.rename(from = here::here(file_name),
            to = here::here('manuscript', 'output', file_name))
}
```

```{r}
library(targets)
library(tidyverse)
library(gt)
library(JakeR)
library(patchwork)
tar_config_set(store = here::here('_targets/'))
tar_load(pooled_dat)
tar_load(meps_iqvia)

```

```{r reporting-functions}
comparePercents <- function(data, prescription, which.first) {
  if (which.first == 'MEPS') {
    data[data$Prescription == prescription, 'MEPS'] - data[data$Prescription == prescription, 'IQVIA']
  } else {
    data[data$Prescription == prescription, 'IQVIA'] - data[data$Prescription == prescription,  'MEPS']
  }
  
}
```

```{r barplot}
meps_iqvia <- meps_iqvia |>
  mutate(insulin_indicator = case_when(
    str_detect(Prescription, 'Insulin -') ~ 'Insulin',
    TRUE ~ 'Not insulin'
  ))

insulin_data <- meps_iqvia |>
  filter(insulin_indicator == 'Insulin')
not_insulin_data <- meps_iqvia |>
  filter(insulin_indicator != 'Insulin')

wide_insulin_data <- insulin_data |>
  filter(stratifier == 'Overall') |>
  mutate(Prescription = str_remove_all(Prescription, 'Insulin - ')) |>
  mutate(Prescription = factor(
    Prescription,
    levels = c(
      'Bolus/Human',
      'Bolus/Analog',
      'Basal/Human',
      'Basal/Analog',
      'Pre-mixed',
      'Pre-mixed/Human',
      'Pre-mixed/Analog',
      'Insulin/Unknown'
    )
  )) |>
  mutate(total = sum(Number, na.rm = TRUE), .by = data) |>
  mutate(Percent = (Number / total) * 100) |> 
  pivot_wider(names_from = 'data',
              values_from = 'Percent',
              id_cols = c(Prescription, stratifier, group))

wide_not_insulin <- not_insulin_data |>
  filter(stratifier == 'Overall') |>
  mutate(total = sum(Number, na.rm = TRUE), .by = data) |>
  mutate(Percent = (Number / total) * 100) |>
  mutate(Prescription = factor(
    Prescription,
    levels = c(
      'All insulins',
      'Biguanides',
      'Sulfonylureas',
      'TZD',
      'DPP-4i',
      'GLP-1RA',
      'SGLT2i',
      'Unknown'
    )
  ))|> 
  pivot_wider(names_from = 'data',
              values_from = 'Percent',
              id_cols = c(Prescription, stratifier, group))

```

```{r heatmap}

wide_meps_iqvia <- meps_iqvia |> 
  filter(stratifier != 'Prescribing medical specialty') |> 
  pivot_wider(names_from = data,
              values_from = c(Number, Percent, PercentLL, PercentUL, Flags),
              id_cols = c(Prescription, stratifier, group)) |> 
  select(-PercentLL_IQVIA, -PercentUL_IQVIA)

total_non_insulin_iqvia <- sum(
  wide_meps_iqvia$Number_IQVIA[wide_meps_iqvia$stratifier == 'Overall' & !str_detect(wide_meps_iqvia$Prescription, 'Insulin -')],
  na.rm = TRUE
)

total_non_insulin_meps <- sum(
  wide_meps_iqvia$Number_MEPS[wide_meps_iqvia$stratifier == 'Overall' & !str_detect(wide_meps_iqvia$Prescription, 'Insulin -')],
  na.rm = TRUE
)

total_insulin_iqvia <- wide_meps_iqvia |> 
  filter(stratifier == 'Overall',
         Prescription == 'All insulins') |> 
  pull(Number_IQVIA)

total_insulin_meps <- wide_meps_iqvia |> 
  filter(stratifier == 'Overall',
         Prescription == 'All insulins') |> 
  pull(Number_MEPS)

wide_meps_iqvia <- wide_meps_iqvia |> 
  mutate(noninsulin_total_IQVIA = total_non_insulin_meps,
         noninsulin_total_MEPS = total_non_insulin_meps,
         insulin_total_IQVIA = total_insulin_iqvia,
         insulin_total_MEPS = total_insulin_meps,
         Percent_IQVIA = case_when(stratifier == 'Overall' & str_detect(Prescription, 'nsulin') ~ 
                                     Number_IQVIA / insulin_total_IQVIA * 100,
                                   stratifier == 'Overall' & !str_detect(Prescription, 'nsulin') ~
                                     Number_IQVIA / noninsulin_total_IQVIA * 100,
                                   TRUE ~ Percent_IQVIA),
         Percent_MEPS = case_when(stratifier == 'Overall' & str_detect(Prescription, 'nsulin') ~ 
                                     Number_MEPS / insulin_total_MEPS * 100,
                                   stratifier == 'Overall' & !str_detect(Prescription, 'nsulin') ~
                                     Number_MEPS / noninsulin_total_MEPS * 100,
                                   TRUE ~ Percent_MEPS)
         ) |> 
  mutate(number_diff = Number_IQVIA - Number_MEPS,
         perc_diff = Percent_IQVIA - Percent_MEPS) |> 
  mutate(across(c(number_diff, perc_diff), ~case_when(!is.na(Flags_MEPS) ~ NA,
                                                      TRUE ~ .x)),
         across(c(PercentLL_MEPS, PercentUL_MEPS), ~as.numeric(.x))) |> 
  mutate(perc_diff_between = between(Percent_IQVIA, PercentLL_MEPS, PercentUL_MEPS))
  
comparison_table <- wide_meps_iqvia |> 
  mutate(number_diff = round(number_diff, digits = -3)) |> 
  pivot_wider(names_from = Prescription,
              values_from = perc_diff,
              id_cols = c(stratifier, group))


heatmap_data <- wide_meps_iqvia |>
  filter(group != 'Unknown') |>
  mutate(Group = case_when(stratifier == 'Overall' ~ 'Overall',
                           TRUE ~ paste0(stratifier, ': ', group))) |> 
  mutate(Group = factor(Group,
                        levels = c('Overall',
                                   'Age: < 10',
                                   'Age: 10 - 19',
                                   'Age: 20 - 39',
                                   'Age: 40 - 59',
                                   'Age: 60 - 74',
                                   'Age: 75+',
                                   'Sex: Female',
                                   'Sex: Male',
                                   'Race: Asian/not Hispanic',
                                   'Race: Black/not Hispanic',
                                   'Race: Hispanic',
                                   'Race: Other Race/not Hispanic',
                                   'Race: White/not Hispanic',
                                   'Education: Less than high school',
                                   'Education: High school',
                                   'Education: Greater than high school',
                                   'Insurance: 3rd party',
                                   'Insurance: Any private',
                                   'Insurance: Medicaid',
                                   'Insurance: Medicare Part D',
                                   'Insurance: Public only',
                                   'Insurance: Cash',
                                   'Insurance: Uninsured',
                                   'Copay: $0',
                                   'Copay: >$0-$10',
                                   'Copay: >$10-$20',
                                   'Copay: >$20-$30',
                                   'Copay: >$30-$75',
                                   'Copay: >$75',
                                   'Poverty: High income (>= 400% of poverty line)',
                                   'Poverty: Middle income (200% - 399% of poverty line)',
                                   'Poverty: Low income (100% - <200% of poverty line)',
                                   'Poverty: Poor (<100% of poverty line)'))) |> 
  mutate(Prescription = str_remove(Prescription, 'Insulin - ')) |> 
  mutate(Prescription = factor(Prescription,
                                            levels = c('Bolus/Human',
                                                       'Bolus/Analog',
                                                       'Basal/Human',
                                                       'Basal/Analog',
                                                       'Pre-mixed',
                                                       'Pre-mixed/Human',
                                                       'Pre-mixed/Analog',
                                                       'Insulin/Unknown',
                                                       'All insulins',
                                                       'Biguanides',
                                                       'Sulfonylureas',
                                                       'TZD',
                                                       'DPP-4i',
                                                       'GLP-1RA',
                                                       'SGLT2i',
                                                       'Unknown'))) |>
  filter(stratifier %nin% c('Race', 'Poverty', 'Education'),
         group %nin% c('< 10', '10 - 19')) |> 
  mutate(color = ifelse(perc_diff < -9 | perc_diff > 9, 'white', 'black'),
         insulin_indicator = case_when(Prescription %in% c('Bolus/Human',
                                                       'Bolus/Analog',
                                                       'Basal/Human',
                                                       'Basal/Analog',
                                                       'Pre-mixed',
                                                       'Pre-mixed/Human',
                                                       'Pre-mixed/Analog',
                                                       'Insulin/Unknown',
                                                       'All insulins') ~ 'insulin',
                                       TRUE ~ 'not insulin'))
heatmap <- heatmap_data |>
  filter(!str_detect(Prescription, 'Unknown')) |> 
  # Alter diff coding a bit without changing labels to get the color scales to look a bit better
  mutate(perc_diff_adj = case_when(perc_diff > 15 ~ 15,
                                   perc_diff < -15 ~ -15,
                                   TRUE ~ perc_diff),
         perc_diff_label = specifyDecimal(perc_diff, 1),
         perc_diff_label = case_when(
           perc_diff_between == FALSE & Prescription == 'All insulins' & stratifier == 'Overall' ~ perc_diff_label,
           perc_diff_between == FALSE ~ paste0(perc_diff_label, '*'),
                                     TRUE ~ perc_diff_label)) |> 
  group_split(insulin_indicator) |>
  map(
    ~ ggplot(.x, aes(
      y = Group,
      x = Prescription,
      fill = perc_diff_adj
    )) +
      geom_tile() +
      colorspace::scale_fill_continuous_diverging(
        palette = 'Purple-Green',
        breaks = c(15, 0,-15),
        labels = c('+\n(NPA higher)',
                   '0',
                   '-\n(MEPS higher)'),
        limits = c(-15, 15)
      ) +
      scale_x_discrete(position = 'top',
                       expand = c(0, 0)) +
      scale_y_discrete(limits = rev,
                       expand = c(0, 0)) +
      geom_text(aes(
        label = perc_diff_label,
        color = color
      ),
      show.legend = FALSE) +
      scale_color_manual(values = c('black', 'white')) +
      labs(fill = 'Difference in percentage estimates\n(NPA % - MEPS %)') +
      theme(legend.position = 'bottom',
            axis.text = element_text(color = 'black'))
  )

wrapped_heatmap <- heatmap[[1]] / heatmap[[2]] +
  plot_layout(axes = 'collect',
              guides = 'collect') +
  plot_annotation(tag_level = 'A') &
  theme(legend.position = 'bottom')
```

**Prescriptions for Antihyperglycemic Medication, United States
2020–2021**

Jacob T. Wittman^1^, PhD, Maribeth C. Lovegrove^2^, MPH, Kai McKeever
Bullard^1^, PhD MPH, Sola Han^1^, PhD PharmD, Andrew I. Geller^2^, MD,
Osatohamwen I. Idubor^1^, MD MHS, Yu Wang^1^, PhD, Stephen R. Benoit^1^,
MD MPH

^1^Division of Diabetes Translation, Centers for Disease Control and
Prevention, Atlanta, GA, United States

^2^Division of Healthcare Quality Promotion, Centers for Disease Control
and Prevention, Atlanta, GA, United States

Corresponding Author: Jacob T. Wittman, ugv4\@cdc.gov, 770-488-3533

Disclaimer: The findings and conclusions in this report are those of the
authors and do not necessarily represent the official position of the
Centers for Disease Control and Prevention.

Keywords:

Word Count: 4075 words

Number of tables and figures: 4

\pagebreak

# Abstract (245 words, limit 250)

**Objective:** To compare data from the IQVIA National Prescription
Audit (NPA) and the Medical Expenditures Panel Survey (MEPS) on the
number of dispensed antihyperglycemic medications.

**Research Design and Methods:** We queried both NPA and MEPS for
antihyperglycemic drug prescriptions in 2020–2021. We report the
estimated total number and percentage of total prescriptions for
sulfonylureas, biguanides, insulin, thiazolidinediones, dipeptidyl
peptidase-4 inhibitors, GLP-1 receptor agonists, SGLT-2 inhibitors, and
antidiabetic combinations. When available from the data sources, we also
stratified by a number of sociodemographic variables.

**Results:** Overall, for most of the classes of medications, estimates
of the percent of total medications for a given medicine class from each
dataset were within five percentage points. However, there was a large
discrepancy in the estimated number of prescriptions of bolus analog
insulins (26.7 million from NPA and 41.6 million from MEPS). Among the
stratifying variables available in both datasets, the largest
discrepancies were seen among the insurance groups, with estimated
percentages differing by as much as 22 percentage points.

**Conclusions:** Overall, estimates between the two datasets are
similar. Differences may be due to differing methodologies and different
sample populations. NPA data covers 93% of retaill pharmacies which
allows for more precise estimates for fills from these pharmacies,
whereas MEPS data comes from household surveys, which may capture a
broader range of pharmacies, although with lower precision than
estimates from NPA. MEPS provides patient-level data and may be more
appropriate for analysis of patient-level factors that impact medication
prescriptions.

\pagebreak

# Introduction

As new medications emerge and prescribing practices evolve,
understanding trends in diabetes medication use is crucial for informing
public health interventions, healthcare policy, and clinical practice.
Two potential sources of data for national surveillance of these
prescriptions are the proprietary International Quality and Value
Institute Advisors (IQVIA) National Prescription Audit (NPA) and the
publicly available Medical Expenditures Panel Survey (MEPS) [@availabl;
@MEPSdocumentation]. Both sources provide data on the number of
prescriptions of different classes of medications. NPA covers
approximately 90% of dispensed prescriptions in the US from pharmacies;
however, information on sociodemographics, behaviors, and other key
characteristics is limited. MEPS is a nationally representative data
source that collects information on prescription fills as self-reported
from households while also providing information on sociodemographic and
behavioral variables that NPA does not provide.

The accuracy of MEPS prescription data has been evaluated by comparing
it to Medicare Part D claims in 2006 and 2007 [@hill2011]. In general,
the concordance between the two datasets was high; when estimates
differed it was usually because respondents underreported the number of
different drugs taken but overreported the number of fills [@hill2011].
These discrepancies were consistent across most sociodemographic groups
[@hill2011]. MEPS has been used to examine disparities in the use of
GLP-1 receptor agonists (GLP-1RA) and SGLT-2 inhibitors (SGLT-2i) among
adults with diabetes based on insurance coverage [@bepo2024]; annual
trends in the continuous use of medications for glucose control, blood
pressure, and lipids among adults with diabetes [@chehal2023]; annual
trends in diabetes medication use patterns across different demographic
and socioeconomic groups [@raval2020]; and annual trends in the use of
insulin and glucose monitoring devices among individuals with diabetes
[@perez-nieves2022].

Meanwhile, NPA has been used in combination with data on internet
searches for SGLT-2i and GLP-1RA to estimate the correlation between
actual prescription rates and the frequency with which these medications
were searched [@dzaye2022]. It has also been used to estimate monthly
prescriptions of these medications and how these more granular time
trends vary by different clinician specialties [@adhikari2022]. The
completeness of NPA in capturing the use of oral anti-cancer medications
was evaluated by linking data from NPA to the Surveillance,
Epidemiology, and End Results (SEER) program from 2006 to 2012
[@tran2020]. They found that only 67% of cases in the SEER program had
matches in the NPA database, although IQVIA coverage was lower during
this time period and oral anti-cancer medications may be provided by
specialty pharmacies not covered by IQVIA [@tran2020]. To date, there
has been no evaluation of how MEPS and NPA compare for reporting the
number and type of prescriptions for antihyperglycemic medications.

This paper examines the estimated number of antihyperglycemic medication
prescriptions in the US during 2020-2021, focusing on a comparative
analysis of NPA and MEPS, as well as evaluating reported differences
between groups. Each data source employs different methodologies and
captures information from potentially non-overlapping segments of the
healthcare system, which may lead to variations in estimates and
insights [@defever2024; @tran2020].

# Research Design and Methods

## Data Sources

We used two nationally representative data sources. First, we queried
drug prescriptions from NPA from 2020–2021 [@iqvia]. NPA provides a
national estimate of prescriptions dispensed, receiving over 3 billion
prescription claims per year [@iqvia; @iqvia2024]. NPA collects
prescription data from 90% of retail, 77% of mail order, and 72%
long-term care pharmacies [@iqvia]. For this analysis, we excluded
long-term care pharmacies so estimates would be comparable to the
non-institutionalized population from MEPS. We also restricted our
analysis to adults (18 years of age or older). Second, we used data from
MEPS for 2020–2021 from the Agency for Healthcare Research and Quality,
accessed through IPUMS [@MEPSdocumentation; @blewett2023ipums]. The MEPS
survey is administered to a nationally representative sample of
households of the non-institutionalized population of the U.S, which are
a subsample of households surveyed in the National Health Interview
Survey. IPUMS provides a harmonized version of MEPS data that
standardizes variable names and coding through time [@blewett].

MEPS prescription medication records provide data on self-reported
prescribed medicine; each record is one household-reported medicine
purchased or otherwise obtained in a year. Participants were asked to
supply the names of prescribed medicines, the reason for the
prescription, the number of times the prescription was obtained, and the
names and locations of pharmacies that filled the prescriptions. MEPS
also collects data from the pharmacies of participants on the strength,
quantity, and days of supply of their medications. Industry experts are
consulted by MEPS to identify outliers. Outliers were identified by
comparing the number of reported prescriptions with the number of days
the participant was in the round for the survey and any identified
outliers imputed. These data were then linked with the person-level data
from the full-year consolidated files to provide demographic data. We
used diabetes status to restrict analyses to drugs purchased by someone
with diabetes. This resulted in a dataset of
`r format(nrow(pooled_dat), big.mark = ',')` records of medication
purchases for
`r format(length(unique(pooled_dat$MEPSID)), big.mark = ',')` adults.

## Analysis

The outcome variable for our analysis is the estimate of the number of
prescriptions purchased of each generic antihyperglycemic medication
among the population of adults with diabetes. For each medication fill
record in MEPS, the name of the medicine is reported as well as the most
commonly used generic name based on the Multum Lexicon database. Generic
medications used in the management of diabetes found in the MEPS include
sulfonylureas, biguanides, insulin, thiazolidinediones (TZD), dipeptidyl
peptidase-4 inhibitors (DPP-4i), GLP-1RA, SGLT2i, and antidiabetic
combinations. Records for antidiabetic combinations in MEPS were split
into two records, one for each of the constitutive drug classes (e.g., a
record for glyburide-metformin was split into a record for sulfonylureas
and a record for biguanides). We used the provided generic drug names
and medicine names to further classify insulin records into basal,
bolus, or pre-mixed as well as human or analog categories. In cases
where the generic drug name conflicted with the provided drug name, the
insulin was classified as "insulin - unknown". Similarly, any
non-insulin medications where provided data conflicted were classified
as "unknown". For the MEPS data,, we estimated the number of
prescriptions purchased for each generic medication previously listed,
accounting for complex survey weights and pooling data for two years
[@MEPSdocumentation]. We also queried NPA for the estimates of
antihyperglycemic medication prescriptions among adults with diabetes
using the same generic drug names.

All estimates were age-adjusted using direct standardization and
standardized to the 2000 U.S. population for ages 20–39, 40–59, 60–64,
65–74, and ≥75 years. We present these estimates overall, as well as
stratified by sex, age, type of health insurance coverage for the
individual that was prescribed the medication, and copay amount.
Additionally, each dataset contains stratifying variables unique to
those data; NPA provides information on the medical specialty of the
provider, while MEPS provides additional sociodemographic data on race
and origin, educational attainment, and ratio of the family income to
poverty line (poverty income ratio). Suppression rules were applied to
results from MEPS to suppress unreliable estimates and to present only
those results that were considered statistically reliable
[@parker2023suppression]. In brief, results were deemed statistically
unreliable if the nominal or effective sample size was less than 10 or
if the relative width of the 95% confidence interval was greater than
160%. We also calculated the percentage point difference between NPA and
MEPS estimates of the percentage of prescriptions for each type of
medication out of the total number of insulin prescriptions and the
total number of prescriptions for diabetes medications, overall and
among different stratifying variables. As a result, we were unable to
make comparisons between all available NPA values and MEPS. Because of
the large sample size of NPA and its coverage of its population, we
treat NPA estimates as census estimates with negligible variance. We
identified percentages that differed between the two datasets if the 95%
confidence interval for the MEPS percentage did not include the NPA
percentage.

# Results

## Insulin

The total number of dispensed insulin prescriptions estimated from NPA
was 84,559,000 and from MEPS was 114,659,000 (Table 1). The estimated
total number of dispensed short-acting analog insulins in the NPA data
was 26,720,000 and this comprised most (93.1%) of the short-acting
insulins dispensed in the NPA data. In contrast, the estimated total
number of dispensed short-acting analog insulins from MEPS was
41,655,000, which was 97.5% of the dispensed short-acting insulins in
the MEPS data. The estimated number of dispensed short-acting human
insulins from NPA was 1,975,000 and from MEPS was 1,054,000. The
estimated number of dispensed analog basal insulins from NPA was
48,704,000 and from MEPS was 58,469,00, while the estimated number of
dispensed human basal insulins from NPA was 2,017,000 and from MEPS was
1,729,000. The estimated number of dispensed pre-mixed insulins from NPA
and MEPS were 5,143,000 and 5,704,000, respectively.

Based on estimates from NPA, females comprised a higher percentage of
people dispensed NPH (58.1%) compared to those dispensed analog basal
insulin (50.4%) (Table 1). Almost 90% of dispensed human short-acting
insulins were for people ≥40 years old while 80% of the dispensed analog
short-acting insulin were for people ≥40 years old. Among human basal
insulin (NPH), 19.6% were dispensed to people age 20–39 years compared
to 8.3% of analog basal insulins dispensed to the same group. Less than
1% of analog basal insulin was dispensed to users who were uninsured.
For NPH, 30.5% of the copays for the dispensed medication were over
\$30. Almost 10% of the dispensed NPH were prescribed by obstetrics and
gynecology specialists while those specialists only prescribed 0.3% of
the dispensed analog basal insulin. Among dispensed NPH, 9.4% were
dispensed to people who were uninsured. Only 6.1% of insulin
prescriptions were for premixed insulins. Estimated total number of
dispensed insulins from each dataset for the stratifying variables are
available in supplementary Table S1.

The largest differences between datasets were seen in the bolus analog
and basal analog categories (Figure 1). For these categories, the MEPS
estimate for the percentage of bolus analog dispensed was
`r specifyDecimal(comparePercents(wide_insulin_data, 'Bolus/Analog', 'MEPS'), 1)`
percentage points higher than the NPA value, while the estimated number
of dispensed bolus analog insulins reported in MEPS was almost double
the value reported in NPA. In contrast, the MEPS estimate for the
percentage of basal analog dispensed was
`r specifyDecimal(abs(comparePercents(wide_insulin_data, 'Basal/Analog', 'MEPS')), 1)`
percentage points lower than NPA. For the remaining insulin groups, MEPS
generally reported similar, although often lower percentages. Based on
the information provided by MEPS,
`r specifyDecimal(pull(filter(wide_insulin_data, Prescription == 'Insulin/Unknown'), MEPS), 1)`%
of insulin dispensed could not be categorized.

Among the stratifying variables, the largest discrepancy between
datasets was observed for the insurance groups (Figure 2A). For example,
the MEPS estimates were between
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Public only' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
and
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Public only' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for all categories of insulin among those with
public insurance, while NPA estimates were as much as
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Any private' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for all insulin dispensed among those with
private insurance.

The magnitude of the differences in estimates between the two data
sources were smaller for the age and sex groups than the insurance
groups (Figure 2A). Estimates among the age groups ranged from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for MEPS (bolus analog dispensed for those aged
60–74) to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for NPA (all insulins estimate for adults aged
40–59). Within the sex groups, only the estimated percentage from NPA
for all insulins for females was
`r specifyDecimal(heatmap_data$perc_diff[heatmap_data$group == 'Female' & heatmap_data$Prescription == 'All insulins'], 1)`
percentage points higher than the same estimate from MEPS.

## Non-insulin medications

In general, NPA and MEPS estimates of the total number of non-insulin
medications were more similar than the estimates for insulins. The
estimated number of biguanides dispensed was 160,080,000 from NPA and
171,995,000 from MEPS, and these numbers comprised 49.3% and 51.1% of
the non-insulin diabetes medications (Table 2). The next largest
estimate was for sulfonylureas, with estimated totals of 53,052,000
(NPA) and 53,439,000 (MEPS). The NPA estimated total TZD dispensed was
11,534,000 while the estimated total from MEPS was 10,559,000. The
estimated total number of DPP-4i dispensed was 28,239,000 and 31,330,000
from NPA and MEPS, respectively. The estimated total number of GLP-1RA
dispensed was 38,221,000 from NPA and 39,797,000 from MEPS. Lastly, the
estimated total number of SLGT2i dispensed was 31,323,000 from NPA and
29,233,000 from MEPS.

Based on NPA data, over ¾ of patients received biguanides for ≤ \$10
(Table 2). Sulfonylureas were the second most commonly dispensed
diabetes medications (16.4%) and 20.0% of patients using that medication
were ≥75 years of age. Thiazolidinediones were the least commonly
dispensed and of those taking that class of medication, 58.3% were male,
64.6% were ≥60 years of age, and 63.7% paid ≤ \$10. DPP-4i were the
second least commonly dispensed medication and of those patients taking
the medication, 15.8% paid \>\$75. A higher percentage of patients
taking GLP1-RA were female (54.4%) and people with private insurance
were dispensed 69.7% of the estimated GLP1-RAs, while 30.1% were
dispensed to people with public insurance. Endocrinologists prescribed
18.7% of the GLP1-RA dispensed, a higher percentage than any of the
other diabetes medication classes. A higher percentage of dispensed
SGLT2i were to male patients (58.1%), 71.2% were to people with private
insurance, 14.2% cost \> \$75, and 14.8% of the dispensed SGLT2i were
from endocrinologists and 2.2% from cardiologists. Estimated totals of
insulins for the different stratifying variables from each dataset are
available in supplementary Table S2.

Non-insulin medication percentages were similar for both NPA and MEPS
across many drug categories (Figures 1B & 2B). As with the insulin
medications,
`r specifyDecimal(pull(filter(wide_not_insulin, Prescription == 'Unknown'), MEPS), 1)`%
of non-insulin medications could not be classified. The largest
discrepancy between the two datasets in the percentage of non-insulin
medications was observed among the insurance groups (Figure 2B). Within
the insurance groups, the discrepancies ranged from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Insurance' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for MEPS (TZD estimate for those on public only
insurance) to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Insurance' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for NPA (TZD estimate for those on any private
insurance). The estimates from NPA were higher for eight of the 14
insurance groups where MEPS estimates were available. Among the copay
groups, MEPS estimates were higher for nine of the 31 copay by
non-insulin medication groups. The differences for copay groups ranged
from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Copay' & heatmap_data$insulin_indicator == 'not insulin' ], na.rm = TRUE)[2]), 1)`
to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Copay' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for NPA and MEPS, respectively. The largest
differences were seen for those with no copay purchasing TZD and those
with a copay between \$0 and \$10 purchasing DPP-4i, which were 16.5 and
15.0 percentage points higher in the MEPS data, respectively.

The age group with the largest difference was the estimate for GLP-1RA
for adults aged 40–59, which had an NPA estimate that was
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher than the MEPS estimate (Figure 2B). The largest
difference where MEPS was the higher estimate was for adults aged 75 or
greater, where the MEPS estimate for sulfonylureas prescriptions was
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher than the NPA estimate. NPA estimates were
higher for biguanides and GLP-1RA among adults aged 40–59. The estimates
for non-insulin medications dispensed within the sex groups followed
similar patterns to the insulin estimates: all of the values for females
were higher in the NPA dataset, while estimates for males were higher in
the MEPS dataset, although only the estimated 95% confidence interval
for GLP-1RA in the MEPS data excluded the NPA estimate.

# Discussion

Our study compared NPA and MEPS as sources of national data on
antihyperlycemic prescriptions during 2020–2021. Both datasets reported
relatively similar estimated numbers and percentages for most
anithyperglycemic prescriptions in the U.S. However, a large difference
in the estimated total number of insulins prescribed was identified,
driven primarily by the discrepancy in the estimated number of
short-acting analog insulin prescriptions. Biguanides were the most
commonly prescribed antihyperglycemic medication, followed by insulins,
as reported by both data sources. Analog insulins in both the
fast-acting and long-acting categories are the most prescribed insulin
types. These results are in-line with findings from other studies. In
one study that used data from MEPS for 2008–2015, biguanides were the
most commonly prescribed class of antihyperglycemic medications overall,
accounting for approximately 59% in 2015 [@raval2020]. A study of
ambulatory insulin treatment using data from IQVIA's National Disease
and Therapeutic Index reported that analog insulins were the most common
insulin treatment, accounting for 92% of the treatment visits during
that time frame [@sarkar2021].

Estimates among insurance groups differed the most between the two
datasets. For all drug classes where MEPS data were available, the
number of prescriptions for those with public only insurance was as much
as 22 percentage points higher, depending on the drug class being
evaluated, than the corresponding NPA estimate, while the opposite was
true for those on any private insurance. These estimates may differ
because NPA may not capture medicines filled through pharmacies
dispensing from hospitals or clinics, especially those associated with
public health insurance, such as Veterans Health Administration
pharmacies, or other specialty pharmacies [@defever2024; @hill2011].
Another study found that NPA only captured 59% of those who were
uninsured, 67% of those with private insurance, and 74% of those on
Medicaid who were in the SEER data on patients using oral anticancer
medications [@tran2020].

Differences between the two data sources may be explained in-part due to
methodological differences in how the data are collected. Data on
prescriptions to MEPS is largely self-reported, and self-reported data
on number of fills and refills from surveys are subject to
misclassification. Respondents tend to over report the number of fills
and under report the total number of medications, although reporting
accuracy was generally high for maintenance drugs taken for chronic
conditions [@hill2011]. Additionally, the accuracy of reporting may vary
among different groups, as high health care consumers have been shown to
underreport prescriptions and respondents with lower literacy are often
inconsistent in reporting [@ritter2001; @schönfeld2021]. However, other
studies have shown that inaccuracies in MEPS tend to be consistent
across different socioeconomic and demographic groups and therefore have
minimal impact on the evaluation of behavioral factors that contribute
to differences in medication usage [@hill2011]. While MEPS does include
data on the number of fills and quantity dispensed, these data may
contain imputed values if complete records are missing or were not
provided, and this may introduce some bias in estimating the total
number of fills.

The populations covered for NPA and MEPS are not identical. NPA has high
coverage of retail pharmacies with relatively small variance estimates,
while MEPS estimates come from a much smaller sample of the
non-institutionalized population, with correspondingly larger variance
estimates. However, MEPS medications are reported by all household
members, so all outpatient dispensing settings should be included. An
investigation of the representativeness and completeness of the NPA
database for studying oral anticancer medication found that only 67% of
cases in the SEER program had matches in the NPA database, and this
varied by subgroup. In particular, NPA only captured 61% of males, 59%
of Non-Hispanic Black patients, and 62% of Hispanic patients in the SEER
program, although this may be due in part to oral anticancer medication
prescriptions being filled by specialty pharmacies [@tran2020]. Lastly,
the proprietary methods NPA uses to produce their estimates are also a
potential source for differences between the two datasets, but it is not
possible to evaluate their impact.

When considering these data sources for us in future analyses, it may be
beneficial to consider them together, when possible, or to consider
which is better suited to the question at hand. Because MEPS data are
available at the patient-level, they can provide insights into
patient-level factors that may influence prescription trends. MEPS also
has the advantage of providing data on race/ethnicity of individuals,
which can allow exploration of racial disparities in diabetes medication
usage. Previous studies have shown that racial disparities in diabetes
medication prescription and use exist, and MEPS can be a valuable tool
in tracking these disparities [@chehal2023; @bepo2024; @eberly2021]. NPA
covers 90% of retail pharmacies in the US and may be better for
identifying national trends and patterns in fills by these pharmacies.
Incorporating both datasets may help triangulate and validate findings.

## Limitations

MEPS data is subject to the limitations of survey data, some of which
have already been described, such as subject recall bias or mistakes in
reporting. Additionally, there has been a steadily declining response
rate for MEPS over time. In 2008, the response rate was 59.3% and this
fell to 27.6% in 2020 and fell further to 21.8% in 2021. Although
non-response can be adjusted for in the design of weights, it is not
possible to know how non-response has biased the survey. We assumed that
all prescriptions reported by NPA were for treatment of diabetes,
although this may not be the case.

# Conclusions

Overall, the estimated number and percentage of antihyperglycemic
medications from both NPA and MEPS are comparable, with the exception of
the large difference in the estimated number of bolus analog insulin
prescriptions. Large differences between the estimates were also found
when the estimates were stratified by insurance group. Each dataset
provides a different set of variables that prescriptions may be
stratified by, and the use of a particular dataset should match the
research questions being asked. Because MEPS provides socioeconomic
demographic data, such as race/ethnicity and highest educational
attainment, it is better suited to addressing questions related to
disparities in the prescription of antihyperglycemic medications. NPA
provides data on the provider specialty, while also providing a larger
coverage of the retail universe that should make it less prone to
fluctuations in sample size. When possible, using both datasets may help
to triangulate and validate findings, providing a more comprehensive
understanding of trends and patterns in antihyperglycemic medication
use.

# Acknowledgements

*Funding:* The authors report no funding sources.

*Conflict of Interest:* The authors of this study declare no conflicts
of interest.

*Author Contributions and Guarantor Statement:*

# References

::: {#refs}
:::

\pagebreak

::: landscape
# Tables

```{r}
tar_load(compare_insulins_gt)
tar_load(compare_not_insulins_gt)
```

```{r}
#| label: tbl-insulin
#| tbl-cap: "Weighted estimates from IQVIA National Prescription Audit and the Medical Expenditure Panel Survey from 2020–2021 for age-adjusted number of prescriptions for different categories of insulin purchased by adults with diabetes in the United States. Hyphens indicate the MEPS value was suppressed, while N/A indicates that the stratifying variable is not available from the data source."
compare_insulins_gt 
gtsave(compare_insulins_gt, here::here('manuscript/output/table1.docx'))
```

\pagebreak

```{r}
#| label: tbl-not_insulin
#| tbl-cap: "Weighted estimates from IQVIA National Prescription Audit and the Medical Expenditure Panel Survey from 2020–2021 for age-adjusted number of prescriptions for non-insulin medications and all insulins used in the management of diabetes purchased by adults with diabetes in the United States. Hyphens indicate the MEPS value was suppressed, while N/A indicates that that stratifying variable is not available from the data source."
compare_not_insulins_gt 
gtsave(compare_not_insulins_gt, here::here('manuscript/output/table2.docx'))
```

\pagebreak
:::

# Figures

Figure 1: Comparison of A) estimated percentage of insulin prescriptions
that come from each type of insulin and B) estimated percentage of all
diabetes prescriptions that come from each type of medication from NPA
and MEPS in the USA, 2020–2021.

```{r}
tar_load(meps_iqvia_plots,
         store = here::here('_targets/'))
fig1 <- meps_iqvia_plots +
  plot_annotation(tag_levels = 'A') &
  theme(plot.caption = element_text(hjust = 0))
ggsave(here::here('manuscript/output/fig1.png'),
       fig1,
       width = 8,
       heigh = 8)
```

Figure 2: Heatmap showing the differences between NPA and MEPS in the
estimated percentages of each type of medication out of A) the total
number of insulin prescriptions and B) the total number of prescriptions
for diabetes medications, overall and among different stratifying
variables in the USA, 2020–2021. Grey cells indicate that the MEPS
estimate was suppressed. Asterisks denote estimates where the 95%
confidence interval for the estimated percentage from MEPS did not
include the estimated percentage from NPA.

```{r}
fig2 <- wrapped_heatmap +
  plot_annotation(tag_levels = 'A') &
  theme(plot.caption = element_text(hjust = 0))
ggsave(here::here('manuscript/output/fig2.png'),
       fig2,
       width = 8,
       height = 8)
```
