---
title: ""
format: 
  docx:
    page-width: 10
    reference-doc: template.docx
editor: source
bibliography: resources/references.bib
citation-abbreviations: resources/medline-abbreviations.json
csl: resources/ama.csl
execute:
  echo: false
  warning: false
project: 
  execute-dir: project
crossref: 
  fig-title: Fig
filters:
  - resources/docx-landscape.lua
---

```{r render-for-colleagues, eval = FALSE, include = FALSE}
#' WHEN READY TO RENDER VERSION FOR SHARING WITH COLLEAGUES USE THIS CODE 
#' TO APPEND A DATE TO THE OUTPUT
save <- FALSE
if (save == TRUE) {
file_name <- paste0(Sys.Date(),"_draft_NPA-MEPS-comparison-manuscript.docx")

quarto::quarto_render('manuscript/QUARTO-Manuscript.qmd', output_file = file_name)

file.rename(from = here::here(file_name),
            to = here::here('manuscript', 'output', file_name))
}
```

```{r}
library(targets)
library(tidyverse)
library(gt)
library(JakeR)
library(patchwork)
tar_config_set(store = here::here('_targets/'))
tar_load(pooled_dat)
tar_load(meps_iqvia)

```

```{r reporting-functions}
comparePercents <- function(data, prescription, which.first) {
  if (which.first == 'MEPS') {
    data[data$Prescription == prescription, 'MEPS'] - data[data$Prescription == prescription, 'IQVIA']
  } else {
    data[data$Prescription == prescription, 'IQVIA'] - data[data$Prescription == prescription,  'MEPS']
  }
  
}
```

```{r barplot}
meps_iqvia <- meps_iqvia |>
  mutate(insulin_indicator = case_when(
    str_detect(Prescription, 'Insulin -') ~ 'Insulin',
    TRUE ~ 'Not insulin'
  ))

insulin_data <- meps_iqvia |>
  filter(insulin_indicator == 'Insulin')
not_insulin_data <- meps_iqvia |>
  filter(insulin_indicator != 'Insulin')

wide_insulin_data <- insulin_data |>
  filter(stratifier == 'Overall') |>
  mutate(Prescription = str_remove_all(Prescription, 'Insulin - ')) |>
  mutate(Prescription = factor(
    Prescription,
    levels = c(
      'Bolus/Human',
      'Bolus/Analog',
      'Basal/Human',
      'Basal/Analog',
      'Pre-mixed',
      'Pre-mixed/Human',
      'Pre-mixed/Analog',
      'Insulin/Unknown'
    )
  )) |>
  mutate(total = sum(Number, na.rm = TRUE), .by = data) |>
  mutate(Percent = (Number / total) * 100) |> 
  pivot_wider(names_from = 'data',
              values_from = 'Percent',
              id_cols = c(Prescription, stratifier, group))

wide_not_insulin <- not_insulin_data |>
  filter(stratifier == 'Overall') |>
  mutate(total = sum(Number, na.rm = TRUE), .by = data) |>
  mutate(Percent = (Number / total) * 100) |>
  mutate(Prescription = factor(
    Prescription,
    levels = c(
      'All insulins',
      'Biguanides',
      'Sulfonylureas',
      'TZD',
      'DPP-4i',
      'GLP-1RA',
      'SGLT2i',
      'Unknown'
    )
  ))|> 
  pivot_wider(names_from = 'data',
              values_from = 'Percent',
              id_cols = c(Prescription, stratifier, group))

```

```{r heatmap}

wide_meps_iqvia <- meps_iqvia |> 
  filter(stratifier != 'Prescribing medical specialty') |> 
  pivot_wider(names_from = data,
              values_from = c(Number, Percent, PercentLL, PercentUL, Flags),
              id_cols = c(Prescription, stratifier, group)) |> 
  select(-PercentLL_IQVIA, -PercentUL_IQVIA)

total_non_insulin_iqvia <- sum(
  wide_meps_iqvia$Number_IQVIA[wide_meps_iqvia$stratifier == 'Overall' & !str_detect(wide_meps_iqvia$Prescription, 'Insulin -')],
  na.rm = TRUE
)

total_non_insulin_meps <- sum(
  wide_meps_iqvia$Number_MEPS[wide_meps_iqvia$stratifier == 'Overall' & !str_detect(wide_meps_iqvia$Prescription, 'Insulin -')],
  na.rm = TRUE
)

total_insulin_iqvia <- wide_meps_iqvia |> 
  filter(stratifier == 'Overall',
         Prescription == 'All insulins') |> 
  pull(Number_IQVIA)

total_insulin_meps <- wide_meps_iqvia |> 
  filter(stratifier == 'Overall',
         Prescription == 'All insulins') |> 
  pull(Number_MEPS)

wide_meps_iqvia <- wide_meps_iqvia |> 
  mutate(noninsulin_total_IQVIA = total_non_insulin_meps,
         noninsulin_total_MEPS = total_non_insulin_meps,
         insulin_total_IQVIA = total_insulin_iqvia,
         insulin_total_MEPS = total_insulin_meps,
         Percent_IQVIA = case_when(stratifier == 'Overall' & str_detect(Prescription, 'nsulin') ~ 
                                     Number_IQVIA / insulin_total_IQVIA * 100,
                                   stratifier == 'Overall' & !str_detect(Prescription, 'nsulin') ~
                                     Number_IQVIA / noninsulin_total_IQVIA * 100,
                                   TRUE ~ Percent_IQVIA),
         Percent_MEPS = case_when(stratifier == 'Overall' & str_detect(Prescription, 'nsulin') ~ 
                                     Number_MEPS / insulin_total_MEPS * 100,
                                   stratifier == 'Overall' & !str_detect(Prescription, 'nsulin') ~
                                     Number_MEPS / noninsulin_total_MEPS * 100,
                                   TRUE ~ Percent_MEPS)
         ) |> 
  mutate(number_diff = Number_IQVIA - Number_MEPS,
         perc_diff = Percent_IQVIA - Percent_MEPS) |> 
  mutate(across(c(number_diff, perc_diff), ~case_when(!is.na(Flags_MEPS) ~ NA,
                                                      TRUE ~ .x)),
         across(c(PercentLL_MEPS, PercentUL_MEPS), ~as.numeric(.x))) |> 
  mutate(perc_diff_between = between(Percent_IQVIA, PercentLL_MEPS, PercentUL_MEPS))
  
comparison_table <- wide_meps_iqvia |> 
  mutate(number_diff = round(number_diff, digits = -3)) |> 
  pivot_wider(names_from = Prescription,
              values_from = perc_diff,
              id_cols = c(stratifier, group))


heatmap_data <- wide_meps_iqvia |>
  filter(group != 'Unknown') |>
  mutate(Group = case_when(stratifier == 'Overall' ~ 'Overall',
                           TRUE ~ paste0(stratifier, ': ', group))) |> 
  mutate(Group = factor(Group,
                        levels = c('Overall',
                                   'Age: < 10',
                                   'Age: 10 - 19',
                                   'Age: 20 - 39',
                                   'Age: 40 - 59',
                                   'Age: 60 - 74',
                                   'Age: 75+',
                                   'Sex: Female',
                                   'Sex: Male',
                                   'Race: Asian/not Hispanic',
                                   'Race: Black/not Hispanic',
                                   'Race: Hispanic',
                                   'Race: Other Race/not Hispanic',
                                   'Race: White/not Hispanic',
                                   'Education: Less than high school',
                                   'Education: High school',
                                   'Education: Greater than high school',
                                   'Insurance: 3rd party',
                                   'Insurance: Any private',
                                   'Insurance: Medicaid',
                                   'Insurance: Medicare Part D',
                                   'Insurance: Public only',
                                   'Insurance: Cash',
                                   'Insurance: Uninsured',
                                   'Copay: 0',
                                   'Copay: >0-10',
                                   'Copay: >10-20',
                                   'Copay: >20-30',
                                   'Copay: >30-75',
                                   'Copay: >75',
                                   'Poverty: High income (>= 400% of poverty line)',
                                   'Poverty: Middle income (200% - 399% of poverty line)',
                                   'Poverty: Low income (100% - <200% of poverty line)',
                                   'Poverty: Poor (<100% of poverty line)'))) |> 
  mutate(Prescription = str_remove(Prescription, 'Insulin - ')) |> 
  mutate(Prescription = factor(Prescription,
                                            levels = c('Bolus/Human',
                                                       'Bolus/Analog',
                                                       'Basal/Human',
                                                       'Basal/Analog',
                                                       'Pre-mixed',
                                                       'Pre-mixed/Human',
                                                       'Pre-mixed/Analog',
                                                       'Insulin/Unknown',
                                                       'All insulins',
                                                       'Biguanides',
                                                       'Sulfonylureas',
                                                       'TZD',
                                                       'DPP-4i',
                                                       'GLP-1RA',
                                                       'SGLT2i',
                                                       'Unknown'))) |>
  filter(stratifier %nin% c('Race', 'Poverty', 'Education'),
         group %nin% c('< 10', '10 - 19')) |> 
  mutate(color = ifelse(perc_diff < -9 | perc_diff > 9, 'white', 'black'),
         insulin_indicator = case_when(Prescription %in% c('Bolus/Human',
                                                       'Bolus/Analog',
                                                       'Basal/Human',
                                                       'Basal/Analog',
                                                       'Pre-mixed',
                                                       'Pre-mixed/Human',
                                                       'Pre-mixed/Analog',
                                                       'Insulin/Unknown',
                                                       'All insulins') ~ 'insulin',
                                       TRUE ~ 'not insulin'))
heatmap <- heatmap_data |>
  filter(!str_detect(Prescription, 'Unknown')) |> 
  # Alter diff coding a bit without changing labels to get the color scales to look a bit better
  mutate(perc_diff_adj = case_when(perc_diff > 15 ~ 15,
                                   perc_diff < -15 ~ -15,
                                   TRUE ~ perc_diff),
         perc_diff_label = specifyDecimal(perc_diff, 1),
         perc_diff_label = case_when(
           perc_diff_between == FALSE & Prescription == 'All insulins' & stratifier == 'Overall' ~ perc_diff_label,
           perc_diff_between == FALSE ~ paste0(perc_diff_label, '*'),
                                     TRUE ~ perc_diff_label)) |> 
  group_split(insulin_indicator) |>
  map(
    ~ ggplot(.x, aes(
      y = Group,
      x = Prescription,
      fill = perc_diff_adj
    )) +
      geom_tile() +
      colorspace::scale_fill_continuous_diverging(
        palette = 'Purple-Green',
        breaks = c(15, 0,-15),
        labels = c('+\n(NPA higher)',
                   '0',
                   '-\n(MEPS higher)'),
        limits = c(-15, 15)
      ) +
      scale_x_discrete(position = 'top',
                       expand = c(0, 0)) +
      scale_y_discrete(limits = rev,
                       expand = c(0, 0)) +
      geom_text(aes(
        label = perc_diff_label,
        color = color
      ),
      show.legend = FALSE) +
      scale_color_manual(values = c('black', 'white')) +
      labs(fill = 'Difference in percentage estimates\n(NPA % - MEPS %)') +
      theme(legend.position = 'bottom',
            axis.text = element_text(color = 'black'))
  )

wrapped_heatmap <- heatmap[[1]] / heatmap[[2]] +
  plot_layout(axes = 'collect',
              guides = 'collect') +
  plot_annotation(tag_level = 'A') &
  theme(legend.position = 'bottom')
```

**Prescriptions for Antihyperglycemic Medication, United States
2020–2021**

Jacob T. Wittman^1^, PhD, Maribeth C. Lovegrove^2^, MPH, Kai McKeever
Bullard^1^, PhD MPH, Sola Han^1^, PhD PharmD, Andrew I. Geller^2^, MD,
Osatohamwen I. Idubor^1^, MD MHS, Yu Wang^1^, PhD, Stephen R. Benoit^1^,
MD MPH

^1^Division of Diabetes Translation, Centers for Disease Control and
Prevention, Atlanta, GA, United States

^2^Division of Healthcare Quality Promotion, Centers for Disease Control
and Prevention, Atlanta, GA, United States

Corresponding Author: Jacob T. Wittman, ugv4\@cdc.gov, 770-488-3533

Disclaimer: The findings and conclusions in this report are those of the
authors and do not necessarily represent the official position of the
Centers for Disease Control and Prevention.

Keywords:

Word Count: 3732 words, 4000 limit

Number of tables and figures: 4

\pagebreak

# Abstract (248 words, limit 250)

**Objective:** To compare data from the IQVIA National Prescription
Audit (NPA) and the Medical Expenditures Panel Survey (MEPS) on the
number of antihyperglycemic medications.

**Research Design and Methods:** We queried both NPA and MEPS for
antihyperglycemic drug prescriptions in 2020–2021. We report the total
number of and percentage of total prescriptions for sulfonylureas,
biguanides, insulin, thiazolidinediones, dipeptidyl peptidase 4
inhibitors, GLP-1 receptor agonists, SGLT-2 inhibitors, and antidiabetic
combinations. When available from the data sources, we also stratified a
number of sociodemographic variables.

**Results:** Overall, for most of the classes of medications, estimates
of the percent of total medications for a given medicine class from each
dataset were within five percentage points. However, there was a large
discrepancy in the raw number of prescriptions of bolus analog insulins
(26.7 million from NPA and 41.6 million from MEPS). Among the
stratifying variables available in both datasets, the largest
discrepancies were seen among the insurance groups, with estimated
percentages differing by as much as 22 percentage points.

**Conclusions:** Overall, estimates between the two datasets are
similar. Differences may be due to differing methodologies and different
sample populations. NPA data covers 90% of retail, mail order, and
long-term care pharmacies which allows for more precise estimates for
fills from these pharmacies, whereas MEPS data comes from household
surveys, which may capture a broader range of pharmacies, although with
lower precision than estimates from NPA. MEPS provides patient-level
data and may be more appropriate for analysis of patient-level factors
that impact medication prescriptions.

\pagebreak

# Introduction

As new medications emerge and prescribing practices evolve,
understanding trends in diabetes medication use is crucial for informing
public health interventions, healthcare policy, and clinical practice.
Two potential sources for data for national surveillance of these
prescriptions are the proprietary International Quality and Value
Institute Advisors (IQVIA) National Prescription Audit (NPA) and the
publicly available Medical Expenditures Panel Survey (MEPS) [@availabl;
@MEPSdocumentation]. Both sources provide data on the number of
prescriptions of different classes of medications. NPA covers
approximately 90% of dispensed prescriptions in the US from pharmacies;
however, information on sociodemographics, behaviors, and other key
characteristics is lacking. MEPS is a nationally representative data
source that collects information on prescription fills as self-reported
from households while also providing information on the sociodemographic
and behavioral variables that NPA does not provide.

The accuracy of MEPS prescription data has been evaluated by comparing
it to Medicare Part D claims in 2006 and 2007 [@hill2011]. In general,
the concordance between the two datasets was high; when they differed it
was because respondents often underreported the number of different
drugs taken but overreported the number of fills [@hill2011]. These
discrepancies were consistent across most sociodemographic groups
[@hill2011]. MEPS has been used to examine disparities in the use of
GLP-1 receptor agonists (GLP-1RA) and SGLT-2 inhibitors (SGLT-2i) among
adults with diabetes based on insurance coverage [@bepo2024]; annual
trends in the continuous use of medications for glucose control, blood
pressure, and lipids among adults with diabetes [@chehal2023]; annual
trends in diabetes medication use patterns across different demographic
and socioeconomic groups [@raval2020]; and annual trends in the use of
insulin and glucose monitoring devices among individuals with diabetes
[@perez-nieves2022].

Meanwhile, NPA has been used in combination with data on internet
searches for SGLT-2i and GLP-1RA to estimate the correlation between
actual prescription rates and the frequency with which these medications
were searched [@dzaye2022]. It has also been used to estimate monthly
prescriptions of these medications and how these more granular time
trends vary by different clinician specialties [@adhikari2022]. The
completeness of NPA in capturing the use of oral anti-cancer medications
was evaluated by linking data from NPA to the Surveillance,
Epidemiology, and End Results Program (SEER) [@tran2020]. They found
that only 67% of cases in the Surveillance, Epidemiology, and End
Results (SEER) program had matches in the NPA database, and this number
varied by race, sex, and insurance status [@tran2020]. To date, there
has been no evaluation of how MEPS and NPA compare for reporting the
number and type of prescriptions for antihyperglycemic medications.

This paper examines the number of antihyperglycemic medication
prescriptions in the US during 2020-2021, focusing on a comparative
analysis of NPA and MEPS, as well as evaluating reported differences
between groups. Each data source employs different methodologies and
captures information from potentially non-overlapping segments of the
healthcare system, which may lead to variations in estimates and
insights [@defever2024; @tran2020].

# Research Design and Methods

## Data Sources

We used two nationally representative data sources. First, we queried
drug prescriptions from NPA from 2020–2021 [@iqvia]. NPA provides a
national estimate of prescriptions dispensed, receiving over 3 billion
prescription claims per year. Over 48,000
<!--# This number comes from  https://www.sciencedirect.com/science/article/pii/S0376871620304622-->
pharmacies are included in the data, representing 90% of all retail
pharmacies. NPA collects prescription data from retail, mail order, and
long-term care pharmacies. For this analysis, we excluded long-term care
pharmacies so estimates would be comparable to the non-institutionalized
population from MEPS. We also restricted our analysis to adults. Second,
we used data from MEPS for 2020–2021 from the Agency for Healthcare
Research and Quality, accessed through IPUMS [@MEPSdocumentation;
@blewett2023ipums]. This survey consists of a nationally representative
sample of households of the non-institutionalized population of the U.S,
which are a subsample of households surveyed in the National Health
Interview Survey. IPUMS provides a harmonized version of MEPS data that
standardizes variable names and coding through time [@blewett].

MEPS prescription medication records provide data on self-reported
prescribed medicine; each record is one household-reported medicine
purchased or otherwise obtained in a year. Participants were asked to
supply the names of prescribed medicine, the reason for the
prescription, the number of times the prescription was obtained, and the
names and locations of pharmacies that filled the prescriptions. MEPS
also collects data from the pharmacies of participants on the strength,
quantity, and days of supply of their medications. Industry experts were
consulted by MEPS to identify outliers. Outliers were identified by
comparing the number of reported prescriptions with the number of days
the participant was in the round for the survey and any identified
outliers imputed. These data were then linked with the person-level data
from the full-year consolidated files to provide demographic data. We
used diabetes status to restrict analyses to drugs purchased by someone
with diabetes. This resulted in a dataset of
`r format(nrow(pooled_dat), big.mark = ',')` records of medication
purchases for
`r format(length(unique(pooled_dat$MEPSID)), big.mark = ',')` adults.

## Analysis

The outcome variable for our analysis is the estimate of the number of
prescriptions purchased of each generic antihyperglycemic medication
among the population of adults with diabetes. For each medication fill
record in MEPS, the name of the medicine is reported as well as the most
commonly used generic name based on the Multum Lexicon database. Generic
medications used in the management of diabetes found in the MEPS include
sulfonylureas, biguanides, insulin, thiazolidinediones (TZD), dipeptidyl
peptidase 4 inhibitors (DPP-4i), GLP-1RA, SGLT2i, and antidiabetic
combinations. Records for antidiabetic combinations in MEPS were split
into two records, one for each of the constitutive drug class (e.g. a
record for glyburide-metformin was split into a record for sulfonylureas
and a record for biguanides). We used the provided generic drug names
and medicine names to further classify insulin records into basal,
bolus, or pre-mixed as well as human or analog categories. In cases
where the generic drug name conflicted with the provided drug name, the
insulin was classified as "insulin - unknown". Similarly, any
non-insulin medications where provided data conflicted were classified
as "unknown". MEPS provides instructions for producing estimates,
accounting for complex survey weights and pooling data for two years, we
estimated the number of prescriptions purchased for each generic
medication previously listed [@MEPSdocumentation]. We also queried NPA
for the same estimates of generic antihyperglycemic medication
prescriptions among adults with diabetes.

All estimates were age-adjusted using direct standardization and
standardized to the 2000 U.S. population for ages 20–39, 40–59, 60–64,
65–74, and ≥75 years. We present these estimates overall, as well as
stratified by sex, age, type of health insurance, and copay amount.
Additionally, each dataset contains stratifying variables unique to
those data; NPA provides information on the medical specialty of the
provider, while MEPS provides additional sociodemographic data on race
and origin, educational attainment, and ratio of the family income to
poverty line (poverty income ratio). Suppression rules were applied to
results from MEPS to suppress unreliable estimates and to present only
those results that were considered statistically reliable
[@parker2023suppression]. We also calculated the percentage point
difference between NPA and MEPS estimates of the percentage of
prescriptions for each type of medication out of the total number of
insulin prescriptions and the total number of prescriptions for diabetes
medications, overall and among different stratifying variables. As a
result, we were unable to make comparisons between all available NPA
values and MEPS. Because of the large sample size of NPA and its
coverage of its population, we treat NPA estimates as census estimates
with negligible variance. We identified percentages that differed
between the two datasets if the 95% confidence interval for the MEPS
percentage did not include the NPA percentage.

# Results

## Insulin

Total prescription count estimates from NPA and MEPS for insulins are in
Table 1. Analog insulins comprised most (93.1%) of the short-acting
insulins prescription claims in the NPA data. Almost 90% of patients
using human short-acting insulin were ≥40 years while 80% of the
patients using analog short-acting insulin were ≥40 years (Table 1).
Estimates from NPA showed analog insulins comprised 96.0% of the basal
insulin prescription claims. Among patients using human basal insulin
(NPH), 19.6% were age 20–39 years compared to 8.3% of those using analog
basal insulins. Females comprised a higher percentage of the NPH users
(58.1%) compared to the analog basal insulin users (50.4%). Almost 10%
of the NPH claims were prescribed by obstetrics and gynecology
specialists while those specialists only prescribed 0.3% of the analog
basal insulin prescriptions. Among NPH users, 9.4% were uninsured. Less
than 1% of analog basal insulin users were uninsured. For NPH users,
30.5% of the patients paid over \$75. Only 6.1% of insulin prescriptions
were for premixed insulins.

Among the different insulin types, the largest differences between
datasets were seen in the bolus analog and basal analog categories
(Figure 1). For these categories, the MEPS estimate for the percentage
of bolus analog prescriptions was
`r specifyDecimal(comparePercents(wide_insulin_data, 'Bolus/Analog', 'MEPS'), 1)`
percentage points higher than the NPA value, while the estimated number
of prescriptions for bolus analog insulins reported in MEPS was almost
double the value reported in NPA. In contrast, the MEPS estimate for the
percentage of basal analog prescriptions was
`r specifyDecimal(abs(comparePercents(wide_insulin_data, 'Basal/Analog', 'MEPS')), 1)`
percentage points lower than NPA. For the remaining insulin groups, MEPS
generally reported similar, although often lower percentages (Figure 1).
Based on the information provided by MEPS,
`r specifyDecimal(pull(filter(wide_insulin_data, Prescription == 'Insulin/Unknown'), MEPS), 1)`%
of insulin prescriptions could not be categorized.

Among insulin prescriptions, the largest discrepancy between datasets
was observed for the insurance groups (Figure 2A). For example, the MEPS
estimates were between
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Public only' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
and
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Public only' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for all categories of insulin prescriptions
among those with public insurance, while NPA estimates were as much as
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$group== 'Any private' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for all insulin prescriptions among those with
private insurance.

The magnitude of the differences in estimates between the two data
sources were smaller for the age and sex groups than the insurance
groups. Estimates among the age groups ranged from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for MEPS (bolus analog prescriptions for those
aged 60–74) to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for NPA (all insulins estimate for adults aged
40–59). Within the sex groups, only the estimated percentage from NPA
for all insulins for females was
`r specifyDecimal(heatmap_data$perc_diff[heatmap_data$group == 'Female' & heatmap_data$Prescription == 'All insulins'], 1)`
percentage points higher than the same estimate from MEPS.

## Non-insulin medications

Total prescription count estimates from NPA and MEPS are available for
non-insulins in Table 2. Estimates from NPA show that almost half
(49.3%) of all non-insulin diabetes medications prescriptions were for
metformin. Over ¾ of patients received the medication for ≤ \$10.
Sulfonylureas were the second most commonly prescribed diabetes
medications (16.4%) and 20.0% of patients using that medication were ≥75
years of age. Thiazolidinediones were the least prescribed and of those
taking that class of medication, 64.7% were ≥60 years of age, 58.3% were
male, and 63.7% paid ≤ \$10. DPP-4i were the second least commonly
prescribed medication and of those patients taking the medication, 15.8%
paid \>\$75. A higher percentage of patients taking GLP1-RA were female
(54.4%) and 18.7% of the prescriptions came from endocrinologists, a
higher percentage than any of the other diabetes medication classes.
Among those prescribed a GLP1-RA, 69.7% had private insurance and 30.1%
had public insurance. A higher percentage of patients taking SGLT2i were
male (58.1%), 14.8% of the prescriptions were from endocrinologists and
2.2% from cardiologists, 71.1% had private insurance, and 14.2% paid \>
\$75.

Non-insulin medication percentages were similar for both NPA and MEPS
across many drug categories (Figures 1B & 2B). As with the insulin
prescriptions,
`r specifyDecimal(pull(filter(wide_not_insulin, Prescription == 'Unknown'), MEPS), 1)`%
of non-insulin medications could not be classified. The largest
discrepancy between the two datasets in the percentage of non-insulin
medications prescriptions was observed among the insurance groups
(Figure 2B). Within the insurance groups, the discrepancies ranged from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Insurance' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for MEPS (TZD estimate for those on public only
insurance) to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Insurance' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher for NPA (TZD estimate for those on any private
insurance). The estimates from NPA were higher for eight of the 14
insurance groups where MEPS estimates were available. Among the copay
groups, MEPS estimates were higher for nine of the 31 copay by
non-insulin medication groups. The differences for copay groups ranged
from
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Copay' & heatmap_data$insulin_indicator == 'not insulin' ], na.rm = TRUE)[2]), 1)`
to
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Copay' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher for NPA and MEPS, respectively. The largest
differences were seen for those with no copay purchasing TZD and those
with a copay between \$0 and \$10 purchasing DPP-4i, which were 16.5 and
15.0 percentage points higher in the MEPS data, respectively.

The age group with the largest difference was the estimate for GLP-1RA
for adults aged 40–59, which had an NPA estimate that was
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[2]), 1)`
percentage points higher than the MEPS estimate. The largest difference
where MEPS was the higher estimate was for adults aged 75 or greater,
where the MEPS estimate for sulfonylureas prescriptions was
`r specifyDecimal(abs(range(heatmap_data$perc_diff[heatmap_data$stratifier == 'Age' & heatmap_data$insulin_indicator == 'not insulin'], na.rm = TRUE)[1]), 1)`
percentage points higher than the NPA estimate. NPA estimates were
higher for biguanides and GLP-1RA among adults aged 40–59. The estimates
for non-insulin medication prescriptions within the sex groups followed
similar patterns to the insulin estimates: all of the values for females
were higher in the NPA dataset, while estimates for males were higher in
the MEPS dataset, although only the estimated 95% confidence interval
for GLP-1RA in the MEPS data excluded the NPA estimate.

# Discussion

Our study compared NPA and MEPS as sources of data on antihyperlycemic
prescriptions during 2020–2021. Both datasets reported relatively
similar estimated numbers and percentages for most anithyperglycemic
prescriptions in the U.S., although there was a large difference in the
estimated total number of insulins prescribed, driven primarily by the
large difference in the estimated number of short-acting analog insulin
prescriptions. Biguanides were the most commonly prescribed medication,
followed by insulins, as reported by both data sources. Analog insulins
in both the fast-acting and long-acting categories are the most
prescribed insulin types. These results are in-line with findings from
other studies. In one study that used data from MEPS for 2008–2015,
biguanides were the most commonly prescribed class of antihyperglycemic
medications overall, accounting for approximately 59% in 2015
[@raval2020]. A study of ambulatory insulin treatment using data from
IQVIA's National Disease and Therapeutic Index reported that analog
insulins were the most common insulin treatment, accounting for 92% of
the treatment visits during that time frame [@sarkar2021].

Estimates among insurance groups differed the most between the two
datasets. For all drug classes where MEPS data were available, the
number of prescriptions for those with public only insurance was as much
as 22 percentage points higher, depending on the drug class being
evaluated, than the corresponding NPA estimate, while the opposite was
true for those on any private insurance. These estimates may differ
because NPA may not capture medicines filled through pharmacies
dispensing from hospitals or clinics, especially those associated with
public health insurance, such as Veterans Health Administration
pharmacies, or other specialty pharmacies [@defever2024; @hill2011].
Another study found that NPA only captured 59% of those who were
uninsured, 67% of those with private insurance, and 74% of those on
Medicaid who were in the SEER data on patients using oral anticancer
medications [@tran2020].

Differences between the two data sources may be explained in-part due to
methodological differences in how the data are collected. Data on
prescriptions to MEPS is largely self-reported, and self-reported data
on number of fills and refills from surveys are subject to
misclassification. Respondents tend to over report the number of fills
and under report the total number of medications, although reporting
accuracy was generally high for maintenance drugs taken for chronic
conditions [@hill2011]. Additionally, the accuracy of reporting may vary
among different groups, as high health care consumers have been shown to
underreport prescriptions and respondents with lower literacy are often
inconsistent in reporting [@ritter2001; @schönfeld2021]. However, other
studies have shown that inaccuracies in MEPS tend to be consistent
across different socioeconomic and demographic groups and therefore have
minimal impact on the evaluation of behavioral factors that contribute
to differences in medication usage [@hill2011]. While MEPS does include
data on the number of fills and quantity dispensed, these data may
contain imputed values if complete records are missing or were not
provided, and this may introduce some bias in estimating the total
number of fills.

The populations covered for NPA and MEPS are not identical. NPA has high
coverage of retail pharmacies with relatively small variance estimates,
while MEPS estimates come from a much smaller sample of the
non-institutionalized population, with correspondingly larger variance
estimates. However, MEPS medications are reported by all household
members, so all outpatient dispensing settings should be included. An
investigation of the representativeness and completeness of the NPA
database for studying oral anticancer medication found that only 67% of
cases in the SEER program had matches in the NPA database, and this
varied by subgroup. In particular, NPA only captured 61% of males, 59%
of Non-Hispanic Black patients, and 62% of Hispanic patients in the SEER
program, although this may be due in part to oral anticancer medication
prescriptions being filled by specialty pharmacies [@tran2020]. Lastly,
the proprietary methods NPA uses to produce their estimates are also a
potential source for differences between the two datasets, but it is not
possible to evaluate how these methods may result in differences.
However, differences in the population are more likely to be the primary
drivers of discrepancies between MEPS and NPA than these proprietary
methods.

It may be beneficial to consider these data sets together, when
possible, or to consider which data source may be more suited to the
question at hand. Because MEPS data is available at the patient-level,
it can provide insights into patient-level factors that may influence
prescription trends. MEPS also has the advantage of providing data on
race/ethnicity of individuals, which can allow exploration of racial
disparities in diabetes medication usage. Previous studies have shown
that racial disparities in diabetes medication prescription and use
exist, and MEPS can be a valuable tool in tracking these disparities
[@chehal2023; @bepo2024; @eberly2021]. NPA covers 90% of retail
pharmacies in the US and may be better for identifying national trends
and patterns in fills by these pharmacies. Incorporating both datasets
may help triangulate and validate findings.

## Limitations

MEPS data is subject to the limitations of survey data, some of which
have already been described, such as subject recall bias or mistakes in
reporting. Additionally, there has been a steadily declining response
rate for MEPS over time. In 2008, the response rate was 59.3% and this
fell to 27.6% in 2020 and fell further to 21.8% in 2021. Although
non-response can be adjusted for in the design of weights, it is not
possible to know how non-response has biased the survey. We assumed that
all prescriptions reported by NPA were for treatment of diabetes, this
may not be the case.

# Conclusions

Overall, the estimated number and percentage of antihyperglycemic
medications from both NPA and MEPS are comparable, with the exception of
the large difference in the estimated number of bolus analog insulin
prescriptions. Large differences between the estimates were also found
when the estimates were stratified by insurance group. Each dataset
provides a different set of variables that prescriptions may be
stratified by, and the use of a particular dataset should match the
questions being asked. Because MEPS provides socioeconomic demographic
data, such as race/ethnicity and highest educational attainment, it is
better suited to addressing questions related to disparities in the
prescription of antihyperglycemic medications. When possible, using both
datasets may help to triangulate and validate findings, providing a more
comprehensive understanding of trends and patterns in antihyperglycemic
medication use.

# Acknowledgements

*Funding:* The authors report no funding sources.

*Conflict of Interest:* The authors of this study declare no conflicts
of interest.

*Author Contributions and Guarantor Statement:*

# References

::: {#refs}
:::

\pagebreak

::: landscape
# Tables

```{r}
tar_load(compare_insulins_gt)
tar_load(compare_not_insulins_gt)
```

```{r}
#| label: tbl-insulin
#| tbl-cap: "Weighted estimates from IQVIA National Prescription Audit and the Medical Expenditure Panel Survey from 2020–2021 for age-adjusted number of prescriptions for different categories of insulin purchased by adults with diabetes in the United States. Hyphens indicate the MEPS value was suppressed, while N/A indicates that that stratifying variable is not available from the data source."
compare_insulins_gt 
gtsave(compare_insulins_gt, here::here('manuscript/output/table1.docx'))
```

\pagebreak

```{r}
#| label: tbl-not_insulin
#| tbl-cap: "Weighted estimates from IQVIA National Prescription Audit and the Medical Expenditure Panel Survey from 2020–2021 for age-adjusted number of prescriptions for non-insulin medications used in the management of diabetes purchased by adults with diabetes in the United States. Hyphens indicate the MEPS value was suppressed, while N/A indicates that that stratifying variable is not available from the data source."
compare_not_insulins_gt 
gtsave(compare_not_insulins_gt, here::here('manuscript/output/table2.docx'))
```

\pagebreak
:::

# Figures

Figure 1: Comparison of A) percentage of insulin prescriptions that come
from each type of insulin and B) percentage of all diabetes
prescriptions that come from each type of medication as estimated from
NPA and MEPS in the USA, 2020–2021.

```{r}
tar_load(meps_iqvia_plots,
         store = here::here('_targets/'))
fig1 <- meps_iqvia_plots +
  plot_annotation(tag_levels = 'A') &
  theme(plot.caption = element_text(hjust = 0))
ggsave(here::here('manuscript/output/fig1.png'),
       fig1,
       width = 8,
       heigh = 8)
```

Figure 2: Heatmap showing the percentage point difference between NPA
and MEPS estimates of the percentage of prescriptions for each type of
medication out of A) the total number of insulin prescriptions and B)
the total number of prescriptions for diabetes medications, overall and
among different stratifying variables in the USA, 2020–2021. Grey cells
indicate that the MEPS estimate was suppressed. Asterisks denote
estimates where the 95% confidence interval for the estimated percentage
from MEPS did not include the estimated percentage from NPA.

```{r}
fig2 <- wrapped_heatmap +
  plot_annotation(tag_levels = 'A') &
  theme(plot.caption = element_text(hjust = 0))
ggsave(here::here('manuscript/output/fig2.png'),
       fig2,
       width = 8,
       height = 8)
```
